---
title: "Death Rate in USA"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    social: [ "twitter", "facebook", "menu"]
    source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(knitr)
library(DT)
library(rpivotTable)
library(ggplot2)
library(plotly)
library(dplyr)
library(openintro)
library(highcharter)
library(ggvis)
library(shiny)
```


```{r}
##file_path <- file.choose()
t <- getwd()

# Construct the file path
file_path <- file.path(t, "death1.csv")

# Read the CSV file
data <- read.csv(file_path)
```

```{r}
mycolors <- c("blue", "#FFC125", "darkgreen", "darkorange")
```

Interactive Data Visualization
=====================================

Row
-------------------------------------

### Causes of Deaths

```{r}
valueBox(paste("Deaths"),
         color = "warning")
```

### Total Number of Deaths over the years

```{r}
valueBox(sum(data$Deaths),
         icon = "fa-user")
```


### 2016

```{r}
ci1 <-  data %>%
       na.omit() %>%
       filter(data$Year=="2016") 
  
valueBox(value= sum(ci1$Deaths),
         icon = 'fa-building')
```

### 2015

```{r}
ci2 <-  data %>%
       na.omit() %>%
       filter(data$Year=="2015") 
  
valueBox(value= sum(ci2$Deaths),
         icon = 'fa-building')
```

### 2014

```{r}
ci3 <-  data %>%
       na.omit() %>%
       filter(data$Year=="2014") 
  
valueBox(value= sum(ci3$Deaths),
         icon = 'fa-building')
```

### 2013

```{r}
ci4 <-  data %>%
       na.omit() %>%
       filter(data$Year=="2013") 
  
valueBox(value= sum(ci4$Deaths),
         icon = 'fa-building')
```

Row
-------------------------------

### Deaths in 1999

```{r}

ci6 <-data %>% filter(data$Year=="1999") %>% group_by(Cause.Name)
p1 <- ci6 %>%
         plot_ly(x = ~Cause.Name,
                 y = ~Deaths,
                 color = "blue",
                 type = 'bar') %>%
         layout(xaxis = list(title = "Causes of Death"),
                yaxis = list(title = 'Count'))
p1
```

### Deaths in 2001

``` {r}
p2 <- data %>%
         group_by(Cause.Name) %>%
         filter(Year=="2001") %>%
         plot_ly(labels = ~Cause.Name,
                 values = ~Deaths,
                 marker = list(colors = mycolors))  %>%
         add_pie(hole = 0.2) %>%
         layout(xaxis = list(zeroline = F,
                             showline = F,
                             showticklabels = F,
                             showgrid = F),
                yaxis = list(zeroline = F,
                             showline = F,
                             showticklabels=F,
                             showgrid=F))

p2
```

### Deaths in States

```{r}
p3 <- plot_ly(data,
              x = ~State,
              y = ~Deaths ,
              text = paste("State:", data$State,
                           "Deaths:",
                           data$Deaths),
              type = "bar") %>%
              layout(xaxis = list(title="States"),
                     yaxis = list(title = "Number of Deaths"))
p3
```

Row
------------------------------------
### Scatter Plot of Deaths Vs death Rate based on age

```{r}
p4 <- plot_ly(data, x=~Deaths) %>%
         add_markers(y = ~Age.adjusted.Death.Rate,
                     text = ~paste("Mileage: ",Age.adjusted.Death.Rate ),
                     showlegend = F) %>%
         add_lines(y = ~fitted(loess(Age.adjusted.Death.Rate ~ Deaths)),
                   name = "Loess Smoother",
                   color = I("#FFC125"),
                   showlegend = T,
                   line = list(width=5)) %>%
         layout(xaxis = list(title = "Number of Deaths"),
                yaxis = list(title = "Age adjusted Death Rate"))
p4
```

### Box Plot of Death Rate 

```{r}
p5<- plot_ly(data, y=~Age.adjusted.Death.Rate, color = ~State, type="box")%>%
         layout(xaxis = list(title = "States"),
                yaxis = list(title = "Age adjusted Death Rate"))
p5
```

Over the years
=======================================

Row
-------------------------------

### Alzeheimers

```{r}
p6 <- data %>%
         group_by(Year) %>%
         filter(Cause.Name=="Alzheimer's disease") %>%
         plot_ly(labels = ~Year,
                 values = ~Deaths,
                 marker = list(colors = mycolors))  %>%
         add_pie(hole = 0.2) %>%
         layout(xaxis = list(zeroline = F,
                             showline = F,
                             showticklabels = F,
                             showgrid = F),
                yaxis = list(zeroline = F,
                             showline = F,
                             showticklabels=F,
                             showgrid=F))
p6
```

### Diabetes

```{r}
p7 <- data %>%
         group_by(Year) %>%
         filter(Cause.Name=="Diabetes") %>%
         plot_ly(labels = ~Year,
                 values = ~Deaths,
                 marker = list(colors = mycolors))  %>%
         add_pie(hole = 0.2) %>%
         layout(xaxis = list(zeroline = F,
                             showline = F,
                             showticklabels = F,
                             showgrid = F),
                yaxis = list(zeroline = F,
                             showline = F,
                             showticklabels=F,
                             showgrid=F))
p7
```

Row
-------------------------------
### Suicide

```{r}
p8 <- data %>%
         group_by(Year) %>%
         filter(Cause.Name=="Suicide") %>%
         plot_ly(labels = ~Year,
                 values = ~Deaths,
                 marker = list(colors = mycolors))  %>%
         add_pie(hole = 0.2) %>%
         layout(xaxis = list(zeroline = F,
                             showline = F,
                             showticklabels = F,
                             showgrid = F),
                yaxis = list(zeroline = F,
                             showline = F,
                             showticklabels=F,
                             showgrid=F))
p8
```

### Unintentional injuries

```{r}
p9 <- data %>%
         group_by(Year) %>%
         filter(Cause.Name=="Unintentional injuries") %>%
         plot_ly(labels = ~Year,
                 values = ~Deaths,
                 marker = list(colors = mycolors))  %>%
         add_pie(hole = 0.2) %>%
         layout(xaxis = list(zeroline = F,
                             showline = F,
                             showticklabels = F,
                             showgrid = F),
                yaxis = list(zeroline = F,
                             showline = F,
                             showticklabels=F,
                             showgrid=F))
p9
```

Map
========================================

### Map

```{r}
ci10<- data %>% filter(Year=="2016") %>% group_by(State.Code) %>% summarize(avg=mean(Deaths))
l <- list(color = toRGB("white"), width = 2)
g <- list(
  scope= "usa",
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
p10 <- plot_geo(ci10, locationmode = 'USA-states') %>% 
                add_trace(
                  z = ~avg, text = ~avg, locations = ~State.Code,
                  color = ~avg, colors = 'Purples'
                ) %>%
                colorbar(title = "Deaths") %>%
                layout(
                  title = 'DEATHS IN VARIOUS STATES<br>(Hover for breakdown)',
                  geo = g
                )
p10
```


Pivot Table {data=vertical-layout=scroll,horizontal-layout=scroll}
=========================================

```{r}
rpivotTable(data,
            aggregatorName = "Count",
            cols= "Age adjusted Death Rate",
            rows = "State",
            rendererName = "Heatmap",
            height="600px",
            overflow="scroll")


```

Data Table
========================================
```{r}
datatable(data,
          caption = "Death Data",
          rownames = T,
          filter = "top",
          options = list(pageLength = 25))
```


Summary {data-orientation=rows} 
===========================================
Row
-----------

### Number of Deaths

```{r}
valueBox(sum(data$Deaths),
         icon = "fa-user" )
```

### Average Death Rate
```{r}
valueBox(round(mean(data$Age.adjusted.Death.Rate),
               digits = 2),
         icon = "fa-area-chart")
```

### Number of Unique Causes of Death

```{r}
valueBox(length(unique(data$Cause.Name)),
         icon = "fa-user")
```

Row
-------------
### Top 5 causes of death
```{r}
ci11<-data %>% filter(Year==2016) %>% group_by(Cause.Name) %>% summarise(countDeaths=sum(Deaths)) %>% arrange(desc(countDeaths))
temp<-ci11 %>% head(5)

datatable(temp,
          caption = "Top 5 causes of Death",
          rownames = T
          )
```

### States with the most Deaths
```{r}
ci11<-data %>% filter(Year==2016) %>% group_by(State) %>% summarise(countDeaths=sum(Age.adjusted.Death.Rate)) %>% arrange(desc(countDeaths))
temp<-ci11 %>% head(10)

datatable(temp,
          caption = "Top 10 states with highest Death Rate",
          rownames = T)
```



About Report
========================================

Created by: ASMITA CHOTANI 

Conclusion

* This is a report on `r sum(data$Deaths)` Deaths.

* This is a report on `r length(unique(data$State))` States.

* This report was generated on `r format(Sys.Date(), format = "%B %d, %Y")`.


